---
title: "Localization processes for Functional Data Analysis"
author: "Antonio Elías"
date: "20/01/2020"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE, fig.align = 'center'}
knitr::opts_chunk$set(echo = TRUE)

```

localFDA
=======

<!-- badges: start -->
[![License](https://img.shields.io/badge/license-GPL%20v3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)
<!-- badges: end -->

## Overview

Software companion for the paper “Localization processes for functional data analysis” (Elías, Antonio, Jiménez, Raúl, and Yukich, Joe, 2020). 

It implements the localization processes and localization distances. 

## Installation 

```{r, eval = FALSE}
#install the package
devtools::install_github("aefdz/localFDA")

#load the package
library(localFDA)
library(ggplot2)
library(patchwork)
library(dplyr)

```

## Test usage
Load the example data and plot it.

```{r}
data(exampleData)
n <- ncol(X)
p <- nrow(X)
t <- as.numeric(rownames(X))

#plot the data set
df_functions <- data.frame(ids = rep(colnames(X), each = p),
                           y = c(X),
                           x = rep(t, n)
                           )

functions_plot <- ggplot(df_functions) + 
  geom_line(aes(x = x, y = y, group = ids, color = ids), color = "black", alpha = 0.25) + xlab("t") + theme(legend.position = "none")


functions_plot
```

### Compute *kth empirical localization processes*
Empirical version of Equation (1) of the paper. For one focal,
```{r}
focal <- "1"

localizarionProcesses_focal <- localizationProcesses(X, focal)$lc
```

Plot localization processes of order $1, 50, 100$ and $200$:
```{r}
df_lc <- data.frame(k = rep(colnames(localizarionProcesses_focal), each = p),
                           y = c(localizarionProcesses_focal),
                           x = rep(t, n-1)
                           )

lc_plots <- list()
ks <- c(1, 50, 100, 200)

for(i in 1:4){
  lc_plots[[i]] <- functions_plot + 
  geom_line(data = filter(df_lc, k == paste0("k=", ks[i])), aes(x = x, y = y, group = k), color = "blue", size = 1) +
    geom_line(data = filter(df_functions, ids == focal), aes(x = x, y = y, group = ids), color = "red", linetype = "dashed", size = 1)+
    ggtitle(paste("k = ", ks[i]))
}

wrap_plots(lc_plots)

```

### Compute *kth empirical localization distances*
Equation (18) of the paper. For one focal,
```{r}
localizationDistances_focal <- localizationDistances(X, focal)

head(localizationDistances_focal)
```

Plot the localization distances:
```{r}
df_ld <- data.frame(k = names(localizationDistances_focal),
                           y = localizationDistances_focal,
                           x = 1:c(n-1)
                           )


ldistances_plot <- ggplot(df_ld, aes(x = x, y = y)) + geom_point() + 
  ggtitle("Localization distances for one focal") + xlab("kth") + ylab("L")

ldistances_plot
```

### Sample $\mu$ and $\sigma$ 

```{r}
localizationStatistics_full <- localizationStatistics(y, robustify = TRUE)

#See the mean and sd estimations for k = 1, 100, 200, 400, 600

localizationStatistics_full$trim_mu[c(1, 100, 200, 400, 600)]
localizationStatistics_full$trim_sd[c(1, 100, 200, 400, 600)]
```

## References

Elías, Antonio, Jiménez, Raúl and Yukich, Joe (2020). Localization processes for functional data analysis (submitted)
